---
  title: "Project 2--Modeling"
author: "Allison Schieber"
date: '`r format(Sys.time(), "%A, %B %d, %Y @ %I:%M %p")`'
output: 
  html_document: 
  theme: yeti
highlight: textmate
code_folding: hide
---
  
  ```{r globalopts, include = FALSE}
library(knitr)
opts_chunk$set(comment = "", message = FALSE, warning = FALSE)
```

## {.tabset}

### **Packages/Data**

Load all packages and datasets here. Use `glimpse` to visualize the structure of each dataset.

Packages Used

```{r}
library(dplyr)
library(ggplot2)
library(tidyverse)
library(moments)
library(moderndive)
```

Fast Food Data

```{r fastfooddata}
fastfood_data <- read.csv(file = url("https://raw.githubusercontent.com/STAT-JET-ASU/Datasets/master/Instructor/fastfood2017.csv"))
glimpse(fastfood_data)
```

Cookout Burgers Data

```{r cookoutdata}
cookout_data <- read.csv(file = url("https://raw.githubusercontent.com/STAT-JET-ASU/Datasets/master/Instructor/cookoutburgers.csv"))
glimpse(cookout_data)
```

Spruce Experiment Data

```{r sprucedata}
spruce_data <- read.csv(file = url("https://raw.githubusercontent.com/STAT-JET-ASU/Datasets/master/Chihara/Spruce.csv"))
glimpse(spruce_data)
```

Monopoly Game Data

```{r monopolydata}
monopoly_data <- read.csv(file = url("https://raw.githubusercontent.com/STAT-JET-ASU/Datasets/master/Instructor/monopolygame.csv"))
glimpse(monopoly_data)
```


### **Problem 1**

Nutritionists recommend against eating fast food because it is high in sodium, saturated fat, trans fat, and cholesterol. Eating too much over a long period of time can lead to health problems such as high blood pressure, heart disease, and obesity. Many fast-food meals contain more than an entire day's worth of recommended calories! Read the description for the [`fastfood2017`](https://stat-jet-asu.github.io/Datasets/InstructorDescriptions/fastfood2017.html) dataset and the `cookoutburgers` dataset (on the same page), then use the data to perform the following analysis.

A) For fast-food burgers, create a correlation matrix for all of the nutritional variables except `servingsize`. Round the entries to two decimal places. Overall, how would you describe the relationships between nutritional components in fast-food burgers  (mostly weak, moderate, strong, varied)? Explain.

```{r burgercorr}
fastfood_matrix <- fastfood_data %>% 
  filter(type == "Burger")
fastfood_matrix <- fastfood_matrix %>% 
  select(calories, totalfat, saturatedfat, transfat, sodium, carbs, sugars, protein)
round(cor(fastfood_matrix), 2)
```

**ANSWER:**

B) Create a scatterplot with a fitted line to show the relationship between calories (x) and total fat (y) for fast-food burgers. Add an informative title and better axis labels than the default variable names, including units of measure.

```{r burgerplot}
ggplot(fastfood_matrix, aes(x = calories, y = totalfat)) +
  geom_point() +
  xlab("Calories (cal)") +
  ylab("Total fat (g)") +
  ggtitle("Relationship between calories and total fat for fast-food burgers") +
  geom_smooth(method = "lm", se = FALSE)
```

C) Fit a linear model to explore the relationship between calories (x) and total fat (y) for fast-food burgers. Display your model results using `get_regression_table()` and `get_regression_summaries()` with Markdown-formatted output. 

```{r burgermodel}
fastfood_linear_model <- lm(totalfat ~ calories, data = fastfood_matrix)
get_regression_table(fastfood_linear_model, print = TRUE)
get_regression_summaries(fastfood_linear_model)
```

D) Use your model and `get_regression_points()` to predict the total fat in the four Cookout burgers. Display the results with Markdown-formatted output. Compare these predictions to the actual total fat values in the `cookoutburgers` dataset. Did your model do a good job of predicting? Explain.

```{r burgerpred}
fastfood_regression_points <- get_regression_points(fastfood_linear_model, newdata = cookout_data)
fastfood_regression_points
```

**ANSWER:**

E) Suppose someone says, "We can interpret R^2^ as the percentage of calories in each burger that comes from fat." Explain why this is incorrect and give a better interpretation of the R^2^ value for this model in context. 

**ANSWER:**


### **Problem 2**

Case Study 1.10 in the MSRR textbook (p. 10) describes an experiment involving manipulations intended to impact the growth of black spruce seedlings. Load the data in [spruce.csv](https://raw.githubusercontent.com/STAT-JET-ASU/Datasets/master/Chihara/Spruce.csv) and perform the following analysis.

A) Once of the independent variables in the experiment was whether or not seedlings received fertilizer. Create a scatterplot with fitted lines to show the relationship between the heights of seedlings at the start of the experiment (x) and their heights after five years (y), with a separate line for the fertilized and not fertilized groups. Add an informative title and better axis labels than the default variable names, including units of measure.

```{r spruceplot}
ggplot(spruce_data, aes(x = Height0, y = Height5, color = Fertilizer)) +
  geom_point() +
  xlab("Height at start of experiment (cm)") + 
  ylab("Height after 5 years (cm)") +
  ggtitle("Height difference in fertilized vs. non-fertilized plants after 5 years") + 
  geom_smooth(method = "lm", se = FALSE)
```

B) Fit a parallel slopes model to show the height of seedlings after five years as a function of starting height and fertilizer category. Display the `summary()` of your linear model. Interpret the values of the coefficients in context, including units.

```{r sprucemodel}
spruce_slopes <- lm(Height5 ~ Height0 + Fertilizer, data = spruce_data)
summary(spruce_slopes)
```

**ANSWER:** For every 1 cm increase of Height0, Height5 is estimated to be 1.15 cm taller. Non-fertilized seedlings are on average 14.7 cm shorter than their fertilized counterparts. The intercept represents the base value added to Height. 

C) Use `get_regression_points()` to extract the fitted values and residuals from the model. Based on this information and other summary measures, do you think this model would do a good job of predicting the heights of other spruce seedlings grown under the same conditions? Explain.

```{r sprucepts}
get_regression_points(spruce_slopes)
```

**ANSWER:** Based on this information, I don't think this model would be useful for predicting the heights of other spruce seedlings grown under the same conditions. The residuals representing the difference between Height5 and Height5_hat are too large to show an accurate prediction and may result in inaccruate data interpretation.


### **Problem 3**

First known as *The Landlord's Game* by its original developers, the game we now know as *Monopoly* was commercially released by Parker Brothers in 1935. There are now over 1000 variations of this classic game in the US and abroad. Read the description for the [`monopoly`](https://stat-jet-asu.github.io/Datasets/InstructorDescriptions/monopoly.html) dataset and use the data to perform the following analysis.

A) Create a scatterplot with fitted lines to show the relationship between the number of spaces from Go (x) and the purchase price of the property (y), with a separate line for each property type. Add an informative title and better axis labels than the default variable names, including units of measure. Would a parallel slopes model be appropriate here? Why or why not? (Note: It is possible to fit parallel slopes models with more than two categories by adding extra terms to the model, so the number of categories is not an issue.)

```{r pricesplot}
ggplot(monopoly_data, aes(x = spaces, y = cost, color = type)) +
  geom_point() +
  xlab("Number (#) of spaces from Go") +
  ylab("Cost ($) of property") +
  ggtitle("Relationship between the price of properties and distance from Go in Monopoly") + 
  geom_smooth(method = "lm", se = FALSE)
```

**ANSWER:** A parallel slopes model would not be appropriate here because the 2 parallel lines only account for 6 out of all the Monopoly properties. This model does not represent the change in price in the majority of the properties (street properties) as their distance from go increases.
  
  B) Fit a linear model that predicts property price as a function of spaces from Go for all the properties in the game (no groups). Display tables of coefficients and summaries for your model. 

```{r allpricemodel}
monopoly_model <- lm(spaces ~ cost, data = monopoly_data)
get_regression_summaries(monopoly_model)
get_regression_table(monopoly_model)
```

C) Fit a linear model that predicts property price as a function of spaces from Go for the street properties only. Display tables of coefficients and summaries for your model. 

```{r streetpricemodel}
monopoly_street_properties <- monopoly_data %>% 
  filter(type == "street")
monopoly_street_model <- lm(spaces ~ cost, data = monopoly_street_properties)
get_regression_summaries(monopoly_street_model)
get_regression_table(monopoly_street_model)
```

D) Discuss the differences between the your models (e.g., equations, R^2^ values, residuals, other relevant measures of fit). Why did taking out the railroads and utilities produce these changes?
  
  **ANSWER:** Railroads and utilities are unique types of properties because they have a zero slope change in price. This means that the distance from go has no effect on them, their price is the same no matter how far from go they are. This can throw off the data represented in the models. Models that included railroads looked to be closer to a linear model because those data points had no change in price with distance from go. In models without railroads and utilities it was much easier to visualize how change in price correlated with distance from go in the majority of properties. 
  
  
  ### Session Info
  
  **Names of Collaborators**: Michael Kim
  
  ```{r}
sessionInfo()
```
